Batched Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design

### 2018-02-03 paper/intro.tex
Paper update

For analog circuits, the design parameters like widths and lengths of
transistors are manually calculated by designers with their experience and the
understanding of the design specifications. However, due to the progress of IC
manufacture technology forecasted by Moore's law, the circuit devices become
more and more complicated, and the parasitic effect of the circuit can no
longer be ignored; on the other hand, the demands for high-performance,
low-power analog circuits is increasing. It is getting much more difficult to
meet the performance and time-to-market requirement with manual designing.
Automated analog circuit design has thus attracted more research interests in
the past decade~\cite{rutenbar2007hierarchical}.

% TODO: Traditional methods using offline model and simulated based methods are
% not good
The analog circuit design automation problems can be formulated as optimization
problems, the aim is to find the optimal vector of design parameters that give
the best figure of merit (FOM). Prior works about analog circuit optimization
include offline model-based approaches
~\cite{colleran2003optimization,daems2003simulation,wang2014enabling} and
simulation-based approaches. The offline model-based methods try to build
accurate global models and then optimize the cheap-to-evaluate models. The
model either comes from designers' manual derivation or regression models like
SVM and ANN. The problem of this approach is that the accurate models are
usually hard to get; for example, in~\cite{wang2014enabling}, 100,000 randomly
simulated points are used to train a sparse polynomial model for an amplifier
circuit with ten design parameters. Simulation-based methods, instead, treat
the circuits as black-box functions that get the FOM values from circuit
simulations, global optimization algorithms are then directly applied to the
black-box functions.  For simulation-based circuit optimization methods,
meta-heuristic algorithms~\cite{phelps2000anaconda, liu2009analog} are widely
used, although these algorithms can explore the whole design space, they have
relatively low convergence rate. When the circuit simulation takes a long time,
both model-based and simulation-based approaches can be very time-consuming.

% TODO: Bayesian optimization is a sequential algorithm, there is a need to
% parallelize it
%
% TODO: cite the two 3D-IC papers of TVLSI
To reduce the number of circuit simulations, the Gaussian process
(GP)~\cite{GPML} model has been introduced as an online surrogate model to
assist the optimization. In~\cite{liu2014gaspad}, GP is combined with
differential evolution algorithm, Recently, the
Bayesian optimization (BO)~\cite{shahriari2016taking} algorithm has been
applied for analog circuit optimization. In~\cite{lyu2017efficient}, the Bayesian optimization algorithm is
firstly introduced for the single- and multi-objective optimization of general analog circuits and has
shown to be more sample-efficient compared to other simulation-based
approaches. In~\cite{wang2017efficient}, the Bayesian optimization algorithm
is combined with adaptive Monte-Carlo sampling to optimize the yield of analog
circuits and static random-access memory (SRAM).

@@

% TODO: Review current methods
% TODO: See how other papers review existing batched BO methods
Batched Bayesian optimization has been a \textcolor{red}{hot} topic in the BO
literature. Related works include the \emph{simulation matching}
method~\cite{azimi2010batch}, the \emph{BUCB} (BLCB for minimization)
method~\cite{desautels2014parallelizing}, the \emph{GP-UCB-PE}
method~\cite{contal2013parallel}, the \emph{local penalization}
method~\cite{gonzalez2016batch}, the \emph{parallel predictive entropy
search}~\cite{shah2015parallel} and the \emph{qKG}
method~\cite{wu2016parallel}.

@@

% TODO: Beiefly introduce my algorithm
All the algorithms mentioned above choose to use one acquisition function, and
except for the simulation matching method and local penalization method which
can select arbitrary acquisition function, other parallelization methods rely
on specific acquisition function, the UCB acquisition function must be used for
BUCB and GP-UCB-PE, and the knowledge gradient method must be used for the qKG
algorithm. As has been stated in~\cite{hoffman2011portfolio}, no one
acquisition function can always outperform other acquisition functions, so
relying on one acquisition function may resulting in poor performance.

In contrast to existing batched bayesian optimization methods, we propose a
different perspective to the parallelization of Bayesian optimization. In this
paper, we propose to parallelize the Bayesian optimization algorithm via a
multi-objective ensemble of acquisition functions. Firstly, in each iteration,
after the GP model is updated, multiple acquisition functions are selected, we
then perform multi-objective optimization to find the \emph{Pareto front} (PF)
of the acquisition functions. The PF represents the best trade-off between
these acquisition functions. When used in sequential mode, the proposed
multi-objective acquisition ensemble (MACE) strategy can be seen as a portfolio
strategy, when batched evaluation is possible, we can sample multiple points on
the PF as there are usually much more points on the PF than the batch size.

The MACE algorithm is tested using several analytical benchmark functions and
two real-world analog circuits, including an operational amplifier with
\textcolor{red}{14} design parameters and a class-E power amplifier with
\textcolor{red}{ten} design parameters. The BLCB method, local penalization
method with expected improvement acquisition function (EI-LP), the qEI methods
and qEI methods are compared with MACE. The proposed MACE method achieved
competitive performance when compared with the listed state-of-the-art
algorithms.

>>>>

For analog circuits, the design parameters like widths and lengths of
transistors are manually calculated by designers with their experience and the
understanding of the design specifications. However, due to the progress of IC
manufacture technology forecasted by Moore's law, the circuit devices become
more and more complicated, and the parasitic effect of the circuit can no
longer be ignored; on the other hand, the demands for high-performance,
low-power analog circuits is increasing. It is getting much more difficult to
meet the performance and time-to-market requirement with manual designing.
Automated analog circuit design has thus attracted much research interests in
the past decade~\cite{rutenbar2007hierarchical}.

% TODO: Traditional methods using offline model and simulated based methods are
% not good
The analog circuit design automation problems can be formulated as optimization
problems, the aim is to find the optimal vector of design parameters that give
the best figure of merit (FOM). Prior works about analog circuit optimization
include offline model-based approaches
~\cite{colleran2003optimization,daems2003simulation,wang2014enabling} and
simulation-based approaches. The offline model-based methods try to build
accurate global models and then optimize the cheap-to-evaluate models. The
model either comes from designers' manual derivation or regression models like
SVM and ANN. The problem of this approach is that the accurate models are
usually hard to get; for example, in~\cite{wang2014enabling}, 100,000 randomly
simulated points are used to train a sparse polynomial model for an amplifier
circuit with ten design parameters.

Simulation-based methods, instead, treat the circuits as black-box functions
that get the FOM values from circuit simulations, global optimization
algorithms are then directly applied to the black-box functions. For
simulation-based circuit optimization methods, meta-heuristic
algorithms~\cite{phelps2000anaconda, liu2009analog} are widely used, although
these algorithms can explore the whole design space, they have relatively low
convergence rate. When the circuit simulation takes a long time, both
model-based and simulation-based approaches can be very time-consuming.

% TODO: Bayesian optimization is a sequential algorithm, there is a need to
% parallelize it
%
% TODO: cite the two 3D-IC papers of TVLSI
To reduce the number of circuit simulations, the Gaussian process
(GP)~\cite{GPML} model has been introduced as an online surrogate model to
assist the optimization. In~\cite{liu2014gaspad}, GP is combined with
differential evolution algorithm; recently, the
Bayesian optimization (BO)~\cite{shahriari2016taking} algorithm has been
applied for analog circuit optimization. In~\cite{lyu2017efficient}, the Bayesian optimization algorithm is
firstly introduced for the single- and multi-objective optimization of general analog circuits and has
shown to be more sample-efficient compared to other simulation-based
approaches. In~\cite{wang2017efficient}, the Bayesian optimization algorithm
is combined with adaptive Monte-Carlo sampling to optimize the yield of analog
circuits and static random-access memory (SRAM).

@@

% TODO: Review current methods
% TODO: See how other papers review existing batched BO methods
Batched Bayesian optimization has been a hot topic in the BO literature.
Related works include the simulation matching (SM)
method~\cite{azimi2010batch}, the BUCB (BLCB for minimization)
method~\cite{desautels2014parallelizing}, the GP-UCB-PE
method~\cite{contal2013parallel}, the local penalization (LP)
method~\cite{gonzalez2016batch}, the parallel predictive entropy (PPES)
search~\cite{shah2015parallel} and the qKG method~\cite{wu2016parallel}.

\textcolor{red}{TODO}{Discuss more details about these algorithms}

@@

% TODO: Beiefly introduce my algorithm
All the algorithms mentioned above choose to use single acquisition function,
and except for the simulation matching method and local penalization method
which can select arbitrary acquisition function, other parallelization methods
rely on a specific acquisition function. The UCB acquisition function must be
used for BUCB and GP-UCB-PE, and the knowledge gradient method must be used for
the qKG algorithm. As is stated in~\cite{hoffman2011portfolio}, no single
acquisition function can always outperform other acquisition functions, relying
on one acquisition function may result in poor performance.

In this paper, we propose to parallelize the Bayesian optimization algorithm
via the multi-objective acquisition ensemble (MACE). Firstly, in each
iteration, after the GP model is updated, multiple acquisition functions are
selected, we then perform multi-objective optimization to find the \emph{Pareto
front} (PF) of the acquisition functions. The PF represents the best trade-off
between these acquisition functions. When batched evaluations are possible, we
can sample multiple points on the optimized PF to accelerate the optimization.

The MACE algorithm is tested using several analytical benchmark functions and
two real-world analog circuits, including an operational amplifier with ten
design parameters and a class-E power amplifier with twelve design parameters.
The BLCB method, local penalization method with expected improvement
acquisition function (EI-LP), qEI methods and qKG methods are compared with
MACE. The proposed MACE method achieved competitive performance when compared
to the listed state-of-the-art algorithms.
